---
title: "R Notebook"
output: html_notebook
---

# R: Data Structures

## Vectors

Vectors are a list-like data structure that can contain any number of items of the same data type. Vector elements are separated by commas, and the entire vector must be wrapped in `c()`

v<-c("Nazrin","Joe","Eva")
v
a<-1:10
a
b<-2.3:8.9
b

### Getting a Vector's Type

You can get the type of the elements inside of a vector using the `typeof()` function
typeof(v)
typeof(a)
typeof(b)

### Getting a Vector's Length

You can get the length of a vector using the `length()` function

length(v)
length(a)
length(b)

### Indexing a Vector

You can get the value at a specific index in your vector by placing the elements position within brackets, ex: `[1]` .

**NOTE: Indexes in R start at 1, NOT 0.**

v[3]
a[2]
b[]

```

### Generating a sequence between two numerical values

seq<-2:26
seq

# Generate a sequence from 5 to 15

se<-5:15
# Generate a sequence with decimals
 k<-2.9:8.9
 

#### Using the seq() function

k<-seq(1,8)
k

You can also create sequences with more customization by using the `seq()` function. With this you can set a "step" value (this is the `by` parameter).


k<-seq(1,7,by=0.3)

# Step from 5 to 15 by increments of .3
 v<-seq(5,15, by=0.3)
 v

### Vector Sorting

You can sort vectors using the `sort()` function

v <- c('D','R','G','A','X')
sort(v)
sort(v, decreasing=TRUE)

### Vector Calculations

v1 <- c(1, 2, 3)
v2 <- c(4, 5, 6)

#### Addition

v1+v2
v

#### Subtraction
v2-v1
v

#### Multiplication

v2*v1

#### Division

v2/v1

### Vector Recycling

If we perform vector calculations on uneven vectors, the elements of the shorter vectors are reused to fill in the remaining length

v1 + c(1, 2)

## Lists



Lists in R are generic objects that contain an ordered collection of objects. Lists can contain a variety of data types (they are heterogeneous). Lists are created with the `list()` function.

nazrin_list <- list(name="Nazrin", age=28, scores=c(95, 91, 99))
nazrin_list


```{r}

```

## Dataframes

Dataframes are generic objects that store tabular data. They are two-dimensional data structures that can store any data-type (they are heterogeneous). Dataframes store lists of vectors of equal length (each vector is, essentially, a row).

Dataframes have the following constraints: 

-   A dataframe must have column names and every row should have a unique name.

-   Each column must have an identical number of items.

-   Each item in a single column must be of the same data type.

-   Different columns may have different data types.

When creating a dataframe, you first create the values for each column, and use these to generate your dataframe. You generate a dataframe using the `data.frame()` function.

df <- data.frame(
  Name = c("John", "Jane", "Doe"),
  Age = c(22, 30, 25),
  Score = c(90, 85, 88)
)
df

### str()

str(df)

You can get the structure of a dataframe using the `str()` function


str(df)

### summary()

You can get a summary of the data inside of a dataframe using the `summary()` function. This returns useful information like the type of each column and statistical information like the mean, median, and mode of numerical columns.

summary(df)

### Retrieving data from a dataframe

You can index specific columns from a dataframe using the dataframe variable, then a `$` , then the name of the column.


df$Name 

df[c(1, 3), c("Name", "Score")] # first and third rows, first and third columns

You can also index specific rows by calling the name of the dataframe, followed by square brackets and the rows you want to retrieve. The square brackets accept two values, the first is the rows you want, the second is the columns.

```{r}
# Get the first two rows and all of the columns

df[1:2, ]
# Get the first and third rows and all of the columns


df[c(1, 3), ] 
# Get the first and third rows and the first and third columns


df[c(1, 3), c("Name", "Score")] 
### Adding New Columns

A new column can be added to a dataframe by indexing the column using the `$` symbol and setting the value to a vector

df$Grade <- c("A", "B", "A")

### Adding New Rows

To add new rows to a dataframe, we need to first create a new dataframe using the same structure as the original, then combine the two using the `rbind()` function

new_row <- data.frame(Name="Smith", Age=28, Score=92, Grade="B")
df <- rbind(df, new_row)


## Matrices

A matrix is a two-dimensional assortment of number data points in columns and rows. In matrices columns run horizontally and rows run vertically. To create a matrix you call the `matrix()` function. You'll have to pass in a vector of data to generate the matrix from, the number of rows you want, and the number of columns you want. Remember, matrices are in column-wise order.


The following parameters are available to you using matrices:

-   **data** is the input vector.


-   **nrow** is the number of rows to be created.

-   **ncol** is the number of columns to be created.

-   **byrow** If TRUE then the input vector elements are arranged by row, if FALSE they are arranged by column. This is false by default.

-   **dimnames** is the names assigned to the rows and columns.

# Matrices
matrix_data <- matrix(1:12, nrow=4, ncol=3)


### Naming Columns and Rows

You can name columns and rows with the optional dimnames parameter

dimnames(matrix_data) <- list(c("row1", "row2", "row3", "row4"), c("col1", "col2", "col3"))

### Accessing Matrix Values

Values in a matrix can be accessed by indexing with square brackets.

element <- matrix_data[2, ] 

# Get the element in the 2nd row and 4th column

```element <- matrix_data[2, 4]

You can also get entire rows and columns.

row1 <- matrix_data[1, ]

### Calculations between two matrices

m1 <- matrix(1:4, nrow=2)
m2 <- matrix(5:8, nrow=2)

#### Addition

m1+m2

#### Subtraction

m2-m1

#### Multiplication

m1*m2

#### Division
m1/m2
## Arrays

Arrays can store data in more than two dimensions.They are n-dimensional. If we create an array with the dimensions of (5,6,6), it will create 6 matrices with 5 rows and 6 columns. Arrays are created with the `array ()` function. The first argument is the vector of data to display, the second is the desired dimensions for your array.

data_vector <- 1:(5 * 6 * 6) 
array_data <- array(data_vector, dim = c(5, 6, 6))
array_data

### Naming Array columns and rows

You can name columns, rows, and matrices using the optional dimnames parameter

dimnames(array_data) <- list(
  Row = paste("Row", 1:5),          # Row names
  Column = paste("Column", 1:6),    # Column names
  Matrix = paste("Matrix", 1:6)     # Matrix names
)

### Accessing Array Elements

You can access array elements with square brackets, then three arguments. The row, the column, then the matrix.


# Get the element in the second row, third column, first matrix

element <- array_data[2, 3, 1]


# Get the second row of all of the matrices

second_row_all_matrices <- array_data[2, , ]

# Get the second column of all of the matrices

third_column_all_matrices <- array_data[, 3, ]

# Get the second matrix

second_matrix <- array_data[, , 2]


## Factors

Factors are used to categorize and store data. Factors are primarily useful when working with categorical data. Factors are useful in data analysis for statistical modeling. Factor data is expected to be a vector of limited, unique values.

Factors are created using the `factor()` function and passing in a vector of data.


# A factor of eye colors

eye_colors <- c("blue", "green", "brown", "blue", "green", "brown")


You can also use the optional levels parameter to specify the order in which the levels should be displayed.

eye_color_factor <- factor(eye_colors)
eye_color_factor

## Homework

1.  Create an array of two 3x3 matrices from the given vectors.


vector1 <- 1:9
vector2 <- 10:18

array_3x3 <- array(c(vector1, vector2), dim = c(3, 3, 2))

array_3x3

2.  In the above matrix that you created, print the value 5 by indexing in to the matrix.

value_5 <- array_3x3[2, 2, 1]  # 5 is in the 2nd row, 2nd column of the 1st matrix
value_5


3.  Print the third row of the first matrix.

third_row_first_matrix <- array_3x3[3, , 1]
third_row_first_matrix

4.  Print the second column of both matrices.

second_column_all_matrices <- array_3x3[, 2, ]
second_column_all_matrices
